On-line Vision-based Engagement Assessment for Interactive Robots

### 2015-08-26 paper.tex
{sig-alternate -> paper}.tex & added a Makefile

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}
\graphicspath{{figs/}}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{HRI}{'15 Chrischurch, New Zealand.}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format}
\author{Fernando Garcia \qquad Séverin Lemaignan \qquad Alexis Jacq \qquad Pierre Dillenbourg\\Computer-Human Interaction in Learning and Instruction Laboratory (CHILI)\\École Polytechnique Fédérale de Lausanne (EPFL)}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

%\numberofauthors{4} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
%\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
%\alignauthor
%Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{1932 Wallamaloo Lane}\\
%       \affaddr{Wallamaloo, New Zealand}\\
%       \email{trovato@corporation.com}
%% 2nd. author
%\alignauthor
%G.K.M. Tobin\titlenote{The secretary disavows
%any knowledge of this author's actions.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{P.O. Box 1212}\\
%       \affaddr{Dublin, Ohio 43017-6221}\\
%       \email{webmaster@marysville-ohio.com}
%% 3rd. author
%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
%one who did all the really hard work.}\\
%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
%       \affaddr{Hekla, Iceland}\\
%       \email{larst@affiliation.org}
%\and  % use '\and' if you need 'another row' of author names
%% 4th. author
%\alignauthor Lawrence P. Leipuner\\
%       \affaddr{Brookhaven Laboratories}\\
%       \affaddr{Brookhaven National Lab}\\
%       \affaddr{P.O. Box 5000}\\
%       \email{lleipuner@researchlabs.org}
%}

@@

We propose a adaptive model based solution to approach head pose estimation that finds correspondences between facial landmarks points and their respective locations in a face model. We use the open-source library dlib\footnote{http://dlib.net} \cite{dlib09}, to detect and extract the skeleton of the subject's face and then, use the obtained landmarkers to calculate the head pose orientation estimation. In this way, we can estimate what the subject was looking at and obtain a 3D scene representation as it is executed as a ROS (Robot Operating System) node that can be seen in figure \ref{rviz}. Compared to intrusive hardware like head mounted displays, eye-trackers, wearables for instance, this approach has a clear advantage of simplicity and integration with robotic systems, but also becomes accurate enough for general purposes in face-to-face interactions. In addition, it allows as to model complex environments, handling different number and positions of participants in the scene.

\begin{figure}
    \centering
    \includegraphics[width=1\columnwidth]{rviz}
    \caption{\small Rviz screenshot with Nao, camera (Nao's feet), face marker with VFOA (gray cone) and blue marker intersecting the field of view.}
    \label{rviz}
\end{figure}

@@

Figure~\ref{realSetup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
robot.

\begin{figure}
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
            nao} robot.  The robot writes on the tactile tablet, the child then
            corrects the robot by directly overwriting its letters on the tablet
            with a stylus. An adult (either a therapist or an experimenter,
            depending on the studies), remains next to the child to guide the work. 
            A second tablet allows to choose demonstrations and the camera captures the subject's face.}
    \label{realSetup}
\end{figure}

A tactile tablet (with a custom application) is used for both the robot and the child to write: during a typical round, the child requests the robot to write something (a single letter, a number or a full word), and push the tablet towards the robot, the robot writes on the tablet by gesturing the writing (but without actually physically touching the tablet), the child then pull back the
tablet, corrects the robot's attempt by writing him/herself on top or next to the robot's writing (see Figure~\ref{fig:diego}), and ``send'' his/her demonstration to the robot by pressing a small button on the tablet. The robot ``learns'' from this demonstration and tries again.

@@

\begin{table}[h!]
	\centering
	\caption{Tracking accuracy in face-to-face scenario}
	\begin{tabular}{|c|c|c|c|c|c|c|c|} \hline
	\small Subject & 1 & 2 & 3 & 4 & 5 & 6 & Avg\\ \hline
	\small Track(\%) & \small78.7 & \small81.1 & \small73.7 & \small75.6 & \small77.1 & \small68.7 & \small75.8 \\ \hline
	\begin{tabular}{@{}c@{}}\small Real vs \\ \small expected(\%)\end{tabular} & \small93.1 & \small  & \small & \small & \small & \small & \small \\ \hline
	\begin{tabular}{@{}c@{}}\small Real vs \\ \small captured(\%)\end{tabular} & \small59.4 & \small & \small & \small & \small & \small & \small \\ \hline
	\begin{tabular}{@{}c@{}}\small Cosine \\ \small affinity \end{tabular} & \small0.72 & \small & \small & \small & \small & \small & \small \\ \hline
	\end{tabular}
	\label{tab:results}
\end{table}

@@

\begin{figure*}
    \centering
    \includegraphics[width=2\columnwidth]{realExpected}
    \caption{\small In blue, real focus accoding to two observer assessment. In cyan, the overlap between the real and the expected focus of attentions. In yellow, the repeated pattern in turn-taking and orange is the experimenter's suggestion to switch to a difficult word}
    \label{fig:realExpected}

@@

@@

@@

% Still under construction
\section{Conclusions}
The rate of attention shift can predict retained learning. If we consider the gaze shift within the feedback episodes, it is possible to distinguish between long and short based on the number of previous repetitions playing with the same word.
%Plot -> Support this with a plot averaging all interactions of all subjects
%Plot -> Time gets reduced at the same time correctness improves

>>>>

\documentclass{sig-alternate}

\graphicspath{{figs/}}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{HRI}{'16 Chrischurch, New Zealand.}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{On-line Vision-based Engagement Assessment for Interactive Robots}
\author{Fernando Garcia \qquad Séverin Lemaignan \qquad Alexis Jacq \qquad Pierre Dillenbourg\\Computer-Human Interaction in Learning and Instruction Laboratory (CHILI)\\École Polytechnique Fédérale de Lausanne (EPFL)}

@@

We propose a adaptive model based solution to approach head pose estimation that finds correspondences between facial landmarks points and their respective locations in a face model. We use the open-source library dlib\footnote{http://dlib.net} \cite{dlib09}, to detect and extract the skeleton of the subject's face and then, use the obtained landmarkers to calculate the head pose orientation estimation (see figure \ref{system}). In this way, we can estimate what the subject was looking at and obtain a 3D scene representation as it is executed as a ROS (Robot Operating System) node that can be seen in figure \ref{rviz}. Compared to intrusive hardware like head mounted displays, eye-trackers, wearables for instance, this approach has a clear advantage of simplicity and integration with robotic systems, but also becomes accurate enough for general purposes in face-to-face interactions. In addition, it allows as to model complex environments, handling different number and positions of participants in the scene.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{system}
    \caption{\small System overview showing the most relevant components involved in the estimation.}
    \label{system}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{rviz_camera}
    \caption{\small Rviz screenshot showing Nao, camera (Nao's feet), face detected with its field of view (gray cone) and blue marker intersecting as current VFOA.}
    \label{rviz}
\end{figure}

@@

Figure~\ref{fig:realSetup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
robot.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
            nao} robot.  The robot writes on the tactile tablet, the child then
            corrects the robot by directly overwriting its letters on the tablet
            with a stylus. An adult (either a therapist or an experimenter,
            depending on the studies), remains next to the child to guide the work. 
            A second tablet allows to choose demonstrations and the camera captures the subject's face.}
    \label{fig:realSetup}
\end{figure}

A tactile tablet (with a custom application) is used for both the robot and the child to write: during a typical round, the child requests the robot to write something (a single letter, a number or a full word), and push the tablet towards the robot, the robot writes on the tablet by gesturing the writing (but without actually physically touching the tablet), the child then pull back the
tablet, corrects the robot's attempt by writing him/herself on top or next to the robot's writing, and ``send'' his/her demonstration to the robot by pressing a small button on the tablet. The robot ``learns'' from this demonstration and tries again.

@@

\begin{table}[h!]
	\centering
	\caption{Tracking accuracy in face-to-face scenario}
	\begin{tabular}{|c|c|c|c|c|c|c|c|} \hline
	\small Subject & 1 & 2 & 3 & 4 & 5 & 6 & Avg\\ \hline
	\small Track(\%) & \small88.2 & \small83.5 & \small90.5 & \small83.1 & \small87.9 & \small85.0 & \small86.4 \\ \hline
	\begin{tabular}{@{}c@{}}\small Real vs \\ \small expected(\%)\end{tabular} & \small92.6 & \small88.8  & \small95.8 & \small93.1 & \small96.5 & \small89.6 & \small92.7 \\ \hline
	\begin{tabular}{@{}c@{}}\small Real vs \\ \small captured(\%)\end{tabular} & \small61.1 & \small85.5 & \small80.8 & \small62.2 & \small63.4 & \small83.9 & \small72.8 \\ \hline
	\begin{tabular}{@{}c@{}}\small Cosine \\ \small affinity \end{tabular} & \small0.73 & \small0.82 & \small0.87 & \small0.73 & \small0.77 & \small0.89 & \small0.8 \\ \hline
	\end{tabular}
	\label{tab:results}
\end{table}

@@

\begin{figure*}
    \centering
    \includegraphics[width=1.8\columnwidth]{realExpected}
    \caption{\small In blue, real VFOA of subject 4 accoding to two observer's assessment. In cyan, the matching between the real and the obtained VFOA. In yellow, the repeated pattern in turn-taking and in orange, the experimenter's suggestion to choose a new word.}
    \label{fig:realExpected}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{heatmap}
    \caption{\small 2D histogram of a subject's head movements.}
    \label{heatmap}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{kmeans}
    \caption{\small K-means clustering of the head movements.}
    \label{kmeans}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[width=1.81\columnwidth]{realCaptured}
    \caption{\small Real VFOA of subject 4 accoding to two observer's assessment. In green, the obtained VFOA using our aproach and in cyan, the mathing between them. Finally, the correspondent numbers with figure \ref{fig:bitmap}.}
    \label{fig:realCaptured}
\end{figure*}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{errorR}
    \caption{\small Linear regression of all subject attempts against the word correctness, measured as the average euclidean distance of each leter in its corresponding eigenspace with respect to a reference leter. The gray region defines the confidence interval.}
    \label{errorR}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{gazeR}
    \caption{\small Linear regression of all subject attempts against the occurrences of a gaze shift from the writing tablet towards the model provided. The gray region defines the confidence interval.}
    \label{gazeR}
\end{figure}

%If less than a 0.5 seconds without tracking -> Precission  
% Maker-less technology decrease the complexity of the system and increase the naturally of the interaction
% Still under construction
\section{Conclusions}
The rate of attention shift can predict retained learning. If we consider the gaze shift within the feedback episodes, it is possible to distinguish between long and short based on the number of previous repetitions playing with the same word.
%Plot -> Support this with a plot averaging all interactions of all subjects -> Done
%Plot -> Time gets reduced at the same time correctness improves -> Did not show any good results
%Plot -> Difference in gaze during feedback from a subject's whose learning is assimilated vs a one is not: The second shows more intermitent tablet-selection gaze directions.
%Diagram dlib/ROS/launch/nodes/params_ini/txt_model_face -
%Need of more subjects to test the hypothesis
